---
title: "CalclulateDHI"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This function calculates the drug homogeneity index for a given MSI data, using the following formula: 
$$DHI = \frac{\frac{\sum{i=1}^{Ng}  \sum{j=Nu}^{Nz} j.P(i,j)}{\sum{i=1}^{Ng} \sum{j=Nu}^ {Nz} P(i,j)}} {Tumor Area} $$
Here, $P$          is the gray level size-zone matrix(GLSZM) <br>
      $Ng$         is number of gray levels/rows in GLSZM. <br>
      $Nz$         is number of size zones/columns in GLSZM. <br>
      $P(i,j)$     is the frequency for particular gray level $i$ occupided size zone $j$.<br>
      $j$          is the absolute size zone value.<br>
      $Tumor Area$ is the size of the tumor tissue.<br>



#### Derive DHI value from synthetic imaging datasets 

```{r cars}
library(HomogenMSI)
data("DHIimages")
drugImg = DHIimages[[20]]
maskImg = DHIimages[[1]]
maskImg[maskImg !=0] =1
print(CalculateDHI(drugImg,maskImg))
print(CalculateDHI(drugImg,maskImg,QuantLevel=0,Nu=5))


```
In similar way, DHI was calculated for the rest of the image.<br>


```{r DHIcal,echo=FALSE}
DHI = list(); maskImg = DHIimages[[1]]; maskImg[maskImg !=0] =1
DHI[[1]] = 1; DHI[[2]] = CalculateDHI(DHIimages[[2]],maskImg); DHI[[3]] = CalculateDHI(DHIimages[[3]],maskImg)
DHI[[4]] = CalculateDHI(DHIimages[[4]],maskImg); DHI[[5]] = CalculateDHI(DHIimages[[5]],maskImg)
DHI[[6]] = CalculateDHI(DHIimages[[6]],maskImg);DHI[[7]] = CalculateDHI(DHIimages[[7]],maskImg)
DHI[[8]] = CalculateDHI(DHIimages[[8]],maskImg);DHI[[9]] = CalculateDHI(DHIimages[[9]],maskImg)
DHI[[10]] = CalculateDHI(DHIimages[[10]],maskImg);DHI[[11]] = CalculateDHI(DHIimages[[11]],maskImg)
DHI[[12]] = CalculateDHI(DHIimages[[12]],maskImg);DHI[[13]] = CalculateDHI(DHIimages[[13]],maskImg)
DHI[[14]] = CalculateDHI(DHIimages[[14]],maskImg);DHI[[15]] = CalculateDHI(DHIimages[[15]],maskImg)
DHI[[16]] = CalculateDHI(DHIimages[[16]],maskImg);DHI[[17]] = CalculateDHI(DHIimages[[17]],maskImg)
DHI[[18]] = CalculateDHI(DHIimages[[18]],maskImg);DHI[[19]] = CalculateDHI(DHIimages[[19]],maskImg)
DHI[[20]] = CalculateDHI(DHIimages[[20]],maskImg)

```

```{r pressure,echo=FALSE}

par(mfrow=c(5,4),tcl=-0.5,  mai=c(0.3,0.3,0.3,0.3))
image(DHIimages[[1]],axes=FALSE, main = paste("DHI =", DHI[[1]],sep = "")); 
image(DHIimages[[2]],axes=FALSE,main = paste("DHI =", DHI[[2]],sep = "") ); 
image(DHIimages[[3]],axes=FALSE, main = paste("DHI =", round(DHI[[3]],4),sep = "")); 
image(DHIimages[[4]],axes=FALSE, main = paste("DHI =", round(DHI[[4]],4),sep = "")); 
image(DHIimages[[5]],axes=FALSE, main = paste("DHI =", round(DHI[[5]],4),sep = ""))
image(DHIimages[[6]],axes=FALSE, main = paste("DHI =", round(DHI[[6]],4),sep = ""));
image(DHIimages[[7]],axes=FALSE, main = paste("DHI =", round(DHI[[7]],4),sep = ""));
image(DHIimages[[8]],axes=FALSE, main = paste("DHI =", round(DHI[[8]],4),sep = ""));
image(DHIimages[[9]],axes=FALSE, main = paste("DHI =", round(DHI[[9]],4),sep = ""));
image(DHIimages[[10]],axes=FALSE, main = paste("DHI =", round(DHI[[10]],4),sep = ""));
image(DHIimages[[11]],axes=FALSE, main = paste("DHI =", round(DHI[[11]],4),sep = ""));
image(DHIimages[[12]],axes=FALSE, main = paste("DHI =", round(DHI[[12]],4),sep = ""));
image(DHIimages[[13]],axes=FALSE, main = paste("DHI =", round(DHI[[13]],4),sep = ""));
image(DHIimages[[14]],axes=FALSE, main = paste("DHI =", round(DHI[[14]],4),sep = ""));
image(DHIimages[[15]],axes=FALSE, main = paste("DHI =", round(DHI[[15]],4),sep = ""))
image(DHIimages[[16]],axes=FALSE, main = paste("DHI =", round(DHI[[16]],4),sep = ""));
image(DHIimages[[17]],axes=FALSE, main = paste("DHI =", round(DHI[[17]],4),sep = ""));
image(DHIimages[[18]],axes=FALSE, main = paste("DHI =", round(DHI[[18]],4),sep = ""));
image(DHIimages[[19]],axes=FALSE, main = paste("DHI =", round(DHI[[19]],4),sep = ""));
image(DHIimages[[20]],axes=FALSE, main = paste("DHI =", round(DHI[[20]],4),sep = ""))

```

Note: our DHI formula will not work for image 1, as only single gray level present.
